Refinement for open automata

### 2021-06-14 2021-Refinement/report.tex
more comments

\begin{exi}[Enable, variable-oriented]
The automaton drawn on the right side of Figure \ref{fig:enable} is \(\OA{\mbrc{»}}{»}{\mbrc{v}}{\mbrc{v \mapsto l}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are:
\begin{align*}
	\OT{»}{»}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{»}{»}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{»}{»}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
It is an alternative implementation of the enable operator which is FH-bisimilar to the previous one \cite{henrio:01299562}. \Ludo{makes no sense: FH bisimlar is still undefined, neither formally nor informally}
Note that in this example \(r\) and \(l\) are a hole names and also closed terms: this is a naming conflict, though it is not ambiguous because \(J \cap \rterms = \emptyset\) always holds \Ludo{the intersection is at the same time empty and non-empty, why not use a different value (or underline them) or just say that they are in fact different names though it is easier to read using r and l for both}. \Ludo{le probleme ici c est qu on a l impression que tu essayes a la fois d etre tres strict et formel et pas vraiment en meme temps, il faut choisir}
The variables that are said to be local to a transition are all the unbound variables in the expressions minus the automaton variables.
For instance \(v\) is a variable of the automaton (used in the guard).
It is not local to the transition.
\Ludo{je comprends ps cette phrase:} In a run, when taking effective transitions (without local variables), its value is not substituted with a closed term.
For instance \nmm{\OT{»}{»}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is an effective transition generated from the second transition but not \nmm{\OT{»}{»}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the forbidden substitution \(v \mapsto l\)).

@@

Note that in the previous paragraph there are two relations:
The first is between states of two specific automata (\(R \subseteq S_1 \times S_2\)), the second is between automata (\(A_1 \leq A_2\)).
Generally the second is implicit because its definition is always \(\exists R \subseteq S_1 \times S_2, P\mpar{R}\) with \(P\) a property, and only talking about \(R\) and \(P\) is sufficient.
However later on in this article we will define more complex relations on automata so we clarify the distinction now.
A relation of second kind will be called ``relation on automata" whereas a relation of the first kind is called ``relation on configurations" as defined below.
In the work on FH-bisimulation \cite{henrio:01055091} such a relation \(R\) was defined as triples in \(S_1 \times S_2 \times \formulas\) where for \(\mpar{s_1, s_2} \in S_1 \times S_2\) there is a unique \(f \in \formulas\) such that \(\mpar{s_1, s_2, f} \in R\).
This is classically the definition of a function from \(S_1 \times S_2\) to \(\formulas\) so we will define it as such.
\begin{defi}[Relation on configurations]
A relation on configurations of \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).

Two states \(s_1 \in S_1, s_2 \in S_2\) with their respective valuations \(\sigma_1: V_1 \to \values, \sigma_2: V_2 \to \values\) are related iff \(\sigma_1 \uplus \sigma_2 \vdash R\mpar{s_1, s_2}\).
\end{defi}
The important properties we consider for the relation on automata are:
\begin{defi} A relation on open automata \(\leq\) is
\defitem \textbf{reflexive} iff \(\forall a, a \leq a\);
\defitem \textbf{transitive} iff \(\forall a\, b\, c, a \leq b \wedge b \leq c \implies a \leq c\);
\defitem \textbf{a preorder} iff it is reflexive and transitive;
\defitem \textbf{correct w.r.t.\@ composition} iff \(\forall a\, b, a\subst{b}{j} \leq a\);
\defitem \textbf{complete w.r.t.\@ composition} iff \(\forall a\, b, a \leq b \implies \exists c, a \cong b\subst{c}{j}\);
\defitem \textbf{context refining for composition} iff \(\forall a\, b\, c, a \leq b \implies a\subst{c}{j} \leq b\subst{c}{j}\);
\defitem \textbf{congruent for composition} iff \(\forall a\, b\, c, a \leq b \implies c\subst{a}{j} \leq c\subst{b}{j}\);
\defitem \textbf{compatible with composition} iff \(\forall a\, b\, c\, d, a \leq b \wedge c \leq d \implies c\subst{a}{j} \leq d\subst{b}{j}\);
\defitem \textbf{compatible with FH-bisimulation} iff \(\forall a\, b\, c, d, a \cong b \wedge c \cong d \wedge a \leq c \implies b \leq d\).
\end{defi}
Reflexivity, transitivity and preorder are classical properties on relations.
Correctness w.r.t.\@ composition means that every composition is considered a refinement.
This property captures the expected behaviour of a refinement relation on a compositionnal structure like the open automata.
Completeness w.r.t.\@ composition means that a refinement corresponds to the left automaton being equivalent to some composition of the right automaton.
This property is useful if we want to characterise the kind of operations that can generate any refined automaton starting from the specification.
We won't have it because extending the domain of the initial valuation will be a refinement which can't be expressed with composition.
Yet a ``completeness w.r.t.\@ some set of operations" is still a strong and interesting property.
Context refinement is the refinement being compatible with composition of the same automaton.
Congruence is the refinement being compatible with being composed in the same automaton.
Compatibility with composition is the conjunction of the two latter (assuming the relation is a preorder).
This property is required on compositionnal structures to be of any use in verification, proofs and for model checking.
Finally compatibility with FH-bisimulation is the refinement not being able to distinguish two FH-bisimilar automata.
Since FH-bisimulation is the pre-existing equivalence relation, we should not introduce an incompatible proof strategy.

The important properties we consider for the relation on states are being a simulation and the prevention of deadlocks as defined in the following definitions.
\begin{defi}[Pre-simulation]
Let \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\).
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a pre-simulation if it satisfies both
\defitem Initial configurations are related: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are related:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
This definition translated to open automata the classical definition of simulation \cite{10.1007/3-540-54430-5_78}.
The last formula means that for every pair of related configurations of the automaton and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the same holes' action (\(J'_1 \cap J'_2\)) matches (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update. % OR variableS
Apart from holes, this is a natural extension of the notion of simulation on LTS, which is the same definition without variables nor holes nor guards.
In FH-bisimulation, holes with the same name will receive the same open automaton when composed, and the resulting open automaton will still be equivalent.
Here we assume that one of the related open automata can be obtained by composing an open automaton in a hole, open automaton which can have holes itself.
This implies that there is no relation between their holes, for instance composing \(A_i\) in \(j_i\) (\(1 \leq i \leq n\)) of \(A_0\) gives the set of holes \(J \defobject \mpar{\biguplus_{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}\) that is neither subset nor superset of \(J_0\).
Yet we still assume that in holes with the same name, the same open automaton will be composed.
Hence only the holes in common have to match their actions.
However it is not sufficient to imply properties on the relation on automatons like being a preorder.
It is only a necessary condition for any relation that will be defined with ``simulation" in its name.

Deadlocks are reachable configurations from which no further transition is possible.
They can arise unpredictably when parallelism is involved (which is the case in our composition) and they break the intended behaviour of a system.
By preventing them we could hope that the refinement preserves some liveness properties.
\begin{defi}[Deadlock prevention, intuitive definition]
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if:
\defitem It is a pre-simulation;
\defitem If from related states, there is a possible transition in the specification, then there is a pair of matching transition:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \to \values, \mpar{\sigma \vdash R\mpar{s_1, s_2}} \implies \\
	\mpar{\exists t_2 \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \sigma \uplus \nu \vdash \fguard{t_2}} \implies \\
	\bigsymb{\exists} \mpar{t_1, t_2} \defobject \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2}, \exists \nu: \fvars{t_1} \uplus \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
Actually, this requirement prevents the presence of deadlock state in the first automaton if there is a related non-deadlock state in the second by contraposition.
On top of that the transition preventing the deadlock has to be an existing transition in the specification.
This is because a valid transition unmatched in the specification is not a deadlock, yet we do not want it.
The direct formulation of the property would also have required the unintuitive quantification on a possibly empty set of transition to state that they are impossible.
The two previous reasons motivated the formulation in contraposed form.

``Deadlock prevention" conflicts with all the properties where composition is involved (for instance ``composition correctness") by disallowing some unwanted cases where an automaton in a hole cannot produce any action that any out-transition expects.
In particular filling a hole with the deadlock automaton (only one state without out-transitions) is not considered a refinement for a deadlock reducing relation.
A way to solve these conflicts is to characterise a composition that do not introduce deadlocks, which will be used instead of the composition introduced earlier.
\begin{defi}[Reachability]
For any open automata \(A \defobject \OAg\), a reachability predicate \(\reach{A}: S \to \rformulas[V]\) is a predicate on states satisfying both
\defitem Inital state is reachable: \(\sigma_0 \vdash \reach{A}\mpar{s_0}\)
\defitem Reachability is preserved across transitions: \nmm{\forall t \defobject \OTg \in T, \fvars{t} \vdash \reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}}
\end{defi}
The reachability predicate is used to characterise the reachable configurations in a run of an automaton.
In fact the role of the predicate is to characterise potentially reachable configurations without having to characterise configurations as the result of a valid path in an automaton.

>>>>

\begin{exi}[Enable, variable-oriented]
The automaton drawn on the right side of Figure \ref{fig:enable} is \(\OA{\mbrc{»}}{»}{\mbrc{v}}{\mbrc{v \mapsto l}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are:
\begin{align*}
	\OT{»}{»}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{»}{»}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{»}{»}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
It is an alternative implementation of the enable operator which is \Ludo{FH-bisimilar}{makes no sense: FH bisimlar is still undefined, neither formally nor informally} to the previous one \cite{henrio:01299562}. 
Note that in this example \(r\) and \(l\) are a hole names and also closed terms: this is a naming conflict, though it is not ambiguous because \(J \cap \rterms = \emptyset\) always holds \Ludo{}{the intersection is at the same time empty and non-empty, why not use a different value (or underline them) or just say that they are in fact different names though it is easier to read using r and l for both}. \Ludo{le probleme ici c est qu on a l impression que tu essayes a la fois d etre tres strict et formel et pas vraiment en meme temps, il faut choisir}
The variables that are said to be local to a transition are all the unbound variables in the expressions minus the automaton variables.
For instance \(v\) is a variable of the automaton (used in the guard).
It is not local to the transition.
\Ludo{In a run,}{je comprends ps cette phrase:}  when taking effective transitions (without local variables), its value is not substituted with a closed term.
For instance \nmm{\OT{»}{»}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is an effective transition generated from the second transition but not \nmm{\OT{»}{»}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the forbidden substitution \(v \mapsto l\)).

@@

Note that in the previous paragraph there are two relations:
The first is between states of two specific automata (\(R \subseteq S_1 \times S_2\)), the second is between automata (\(A_1 \leq A_2\)).
\Ludo
{Generally the second is implicit because its definition is always \(\exists R \subseteq S_1 \times S_2, P\mpar{R}\) with \(P\) a property, and only talking about \(R\) and \(P\) is sufficient.}{je comprends pas cette phrase}
However later on in this article we will define more complex relations on automata so we clarify the distinction now.
A relation of second kind will be called ``relation on automata" whereas a relation of the first kind is called ``relation on configurations" as defined below.
In the work on FH-bisimulation \cite{henrio:01055091} such a relation \(R\) was defined as triples in \(S_1 \times S_2 \times \formulas\) where for \(\mpar{s_1, s_2} \in S_1 \times S_2\) there is a unique \(f \in \formulas\) such that \(\mpar{s_1, s_2, f} \in R\).
This is classically the definition of a function from \(S_1 \times S_2\) to \(\formulas\) so we will define it as such.
\begin{defi}[Relation on configurations]
A relation on configurations of \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).

Two states \(s_1 \in S_1, s_2 \in S_2\) with their respective valuations \(\sigma_1: V_1 \to \values, \sigma_2: V_2 \to \values\) are related iff \(\sigma_1 \uplus \sigma_2 \vdash R\mpar{s_1, s_2}\).
\end{defi}
The important properties we consider for the relation on automata are:
\begin{defi} A relation on open automata \(\leq\) is
\defitem \textbf{reflexive} iff \(\forall a, a \leq a\);
\defitem \textbf{transitive} iff \(\forall a\, b\, c, a \leq b \wedge b \leq c \implies a \leq c\);
\defitem \textbf{a preorder} iff it is reflexive and transitive;
\defitem \textbf{correct w.r.t.\@ composition} iff \(\forall a\, b, a\subst{b}{j} \leq a\);
\defitem \textbf{complete w.r.t.\@ composition} iff \(\forall a\, b, a \leq b \implies \exists c, a \cong b\subst{c}{j}\);
\defitem \textbf{context refining for composition} iff \(\forall a\, b\, c, a \leq b \implies a\subst{c}{j} \leq b\subst{c}{j}\);
\defitem \textbf{congruent for composition} iff \(\forall a\, b\, c, a \leq b \implies c\subst{a}{j} \leq c\subst{b}{j}\);
\defitem \textbf{compatible with composition} iff \(\forall a\, b\, c\, d, a \leq b \wedge c \leq d \implies c\subst{a}{j} \leq d\subst{b}{j}\);
\defitem \textbf{compatible with FH-bisimulation} iff \(\forall a\, b\, c, d, a \cong b \wedge c \cong d \wedge a \leq c \implies b \leq d\).
\end{defi}
Reflexivity, transitivity and preorder are classical properties on relations.
Correctness w.r.t.\@ composition means that every composition is considered a refinement. \Ludo{problem}{ you have not yet said that your relation was a refinement!}
This property captures the expected behaviour of a refinement relation on a compositional structure like the open automata.
Completeness w.r.t.\@ composition means that a refinement corresponds to the left automaton being equivalent to some composition of the right automaton.
This property is useful if we want to characterise the kind of operations that can generate any refined automaton starting from the specification.
\Ludo{}{I do not understand this: what is the domain of the valuation?}
We won't have it because extending the domain of the initial valuation will be a refinement which can't be expressed with composition.
Yet a ``completeness w.r.t.\@ some set of operations" is still a strong and interesting property.
Context refinement is the refinement being compatible with composition of the same automaton.
Congruence is the refinement being compatible with being composed in the same automaton.
Compatibility with composition is the conjunction of the two latter (assuming the relation is a preorder).
This property is required on compositional structures to be of any use in verification, proofs and for model checking.
Finally compatibility with FH-bisimulation is the refinement not being able to distinguish two FH-bisimilar automata.
Since FH-bisimulation is the pre-existing equivalence relation, we should not introduce an incompatible proof strategy.

\Ludo{ }{I do not understand the english + too many important properties you should give a hierarchy or at least find a better transition}
The important properties we consider for the relation on states are being a simulation and the prevention of deadlocks as defined in the following definitions.
\begin{defi}[Pre-simulation]
Let \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\).
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a pre-simulation if it satisfies both
\defitem Initial configurations are related: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are related:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
This definition translated to open automata the classical definition of simulation \cite{10.1007/3-540-54430-5_78}.
The last formula means that for every pair of related configurations of the automaton and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the same holes' action (\(J'_1 \cap J'_2\)) matches (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update. % OR variableS
Apart from holes, this is a natural extension of the notion of simulation on LTS, which is the same definition without variables nor holes nor guards.
In FH-bisimulation, holes with the same name will receive the same open automaton when composed, and the resulting open automaton will still be equivalent.
Here we assume that one of the related open automata can be obtained by composing an open automaton in a hole, open automaton which can have holes itself.
This implies that there is no relation between \Ludo{their}{whose? refine} holes, for instance composing \(A_i\) in \(j_i\) (\(1 \leq i \leq n\)) of \(A_0\) gives the set of holes \(J \defobject \mpar{\biguplus_{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}\) that is neither subset nor superset of \(J_0\).
Yet we still assume that in holes with the same name, the same open automaton will be composed.
Hence only the holes in common have to match their actions.
However it is not sufficient to imply properties on the relation on automatons like being a preorder.
It is only a necessary condition that we put on any relation that we call a ``simulation'' in the following.

Deadlocks are reachable configurations from which no further transition is possible.
They can arise unpredictably when parallelism is involved (which is the case in our composition) and they generally break the intended behaviour of a system.
By preventing them we \Ludo{could}{This seems to say that it is useless to prevent them} hope that the refinement preserves some liveness properties.
\begin{defi}[Deadlock prevention, intuitive definition]
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if:
\defitem It is a pre-simulation;
\defitem If from related states, there is a possible transition in the \Ludo{specification}{I do not believe specification has been defined nor implementation}, then there is \Ludo{a pair of matching transition}{A matching transition in the IMPLEMENTATION}:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \to \values, \mpar{\sigma \vdash R\mpar{s_1, s_2}} \implies \\
	\mpar{\exists t_2 \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \sigma \uplus \nu \vdash \fguard{t_2}} \implies \\
	\bigsymb{\exists} \mpar{t_1, t_2} \defobject \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2}, \exists \nu: \fvars{t_1} \uplus \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
Actually, this requirement prevents the presence of deadlock state in the first automaton if there is a related non-deadlock state in the second by \Ludo{contraposition}{The contraposition is too tricky for me: are you sure?}.
Additionally, the transition preventing \Ludo{the deadlock}{in the implementation? how can it be the same?why is there a single one? this sentence makes no sense to me} has to be an existing transition in the specification.
This is because a valid transition unmatched in the specification \Ludo{is not a deadlock}{how can a transition be a deadlock? I am completely lost}, yet \Ludo{we do not want it}{too informal}.
The \Ludo{direct formulation}{usually you say ``the direct'' when the property seems to be defined negatively which is syntactically not the case here: you should explain or remove the remark} of the property would also have required the unintuitive quantification on a possibly empty set of transition to state that they are impossible.
The two previous reasons motivated the formulation in contraposed form.

``Deadlock prevention" conflicts with all the properties where composition is involved (for instance ``composition correctness") by disallowing some unwanted cases where an automaton in a hole cannot produce any action that any out-transition expects.
In particular filling a hole with the deadlock automaton (only one state without out-transitions) is not considered a refinement for a deadlock reducing relation.

\Ludo{Here or before you should explain exactly WHAT YOU WANT: I do not know which are the required property which you will achieve which are incompatible, or state that you will now investigate each property and state whether it is desired / achievable/ compatible with the others: without this analysis the reader is lost}{}

A way to solve \Ludo{these conflicts}{which conflict? can be clarified by answering the previous remark} is to characterise a composition that do not introduce deadlocks, which will be used instead of the composition introduced earlier.
\begin{defi}[Reachability]
For any open automata \(A \defobject \OAg\), a reachability predicate \(\reach{A}: S \to \rformulas[V]\) is a predicate on states satisfying both
\defitem Inital state is reachable: \(\sigma_0 \vdash \reach{A}\mpar{s_0}\)
\defitem Reachability is preserved across transitions: \nmm{\forall t \defobject \OTg \in T, \fvars{t} \vdash \reach{A}\mpar{s} \wedge g \implies \reach{A}\mpar{s'}\psubst{\psi}}
\end{defi}
The reachability predicate is used to characterise the reachable configurations in a run of an automaton.
In fact the role of the predicate is to characterise potentially reachable configurations without having to characterise configurations as the result of a valid path in an automaton.
